Resilience and Home-Space for WSTS \thanks{This work was partly done while the authors were supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}

### 2023-10-24 VMCAI-revised-version/final.tex
up

\mathieu{Donner des exemples dans cette classe.}
\alain{je pense aux LCM avec insertions, elle  devrait avoir toutes les propriétés mais faut vérifier}

@@

This all changes when one considers VASS (having control-states). For a VASS $V = (Q,T)$, a set $\Bad = \mathop{\uparrow} \Bad$, for all $q \in Q$, either there is an element of the basis of $\Bad$ with state $q$ \----
then $q(\textbf{v})$ with $\textbf{v} > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none. If there is none, then 
$\mathop{\uparrow} q(\textbf{0})$
 is in the complement of $\Bad$, i.e. $\Safe$. One can then exhibit upward-closed subsets of 
 $\Safe$ and compute basis for the sets of elements from which they are reachable in at most $k$ steps. 
Substracting these predecessor from $\Bad$ yields either a finite number of elements from which one has to check $\Safe$ is reachable in at most $k$ steps, either an infinite number of elements of which there is one which cannot reach $\Safe$ in at most $k$ steps \-- for much the same reasons {\sc Bounded resilience} and {\sc $k$-resilience} never hold for VAS when $\Safe = \mathop{\downarrow} \Safe$ and $\Bad = \mathop{\uparrow} \Bad$. This lead to a decision procedure
 which lead to the following decidability result:
\begin{theorem}
{\sc $k$-resilience }  and {\sc Bounded resilience} are decidable for VASS when 
$\Safe = \mathop{\downarrow} \Safe$.
\end{theorem}

@@

@@

\section{Thm 9}

\mathieu{mettre ici le détail du Theo 9}

\section{More details on VASS}

@@

>>>>

% \mathieu{Donner des exemples dans cette classe.}
% \alain{je pense aux LCM avec insertions, elle  devrait avoir toutes les propriétés mais faut vérifier}

$\post$-effective, $\post$-ideal-effective WSTS, strong downward compatibility and upward compatibility include for instance
Lossy Channel Systems with insertions~\cite{DBLP:journals/iandc/AbdullaJ96, DBLP:journals/iandc/CeceFI96}.

@@

This all changes when one considers VASS (having control-states). For a VASS $V = (Q,T)$, a set $\Bad = \mathop{\uparrow} \Bad$, for all $q \in Q$, either there is an element of the basis of $\Bad$ with state $q$ \----
then $q(\textbf{v})$ with $\textbf{v} > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none. If there is none, then 
$\mathop{\uparrow} q(\textbf{0})$
 is in the complement of $\Bad$, i.e. $\Safe$. One can then exhibit upward-closed subsets of 
 $\Safe$ and compute basis for the sets of elements from which they are reachable in at most $k$ steps. 
Substracting these predecessor from $\Bad$ yields either a finite number of elements from which one has to check $\Safe$ is reachable in at most $k$ steps, either an infinite number of elements of which there is one which cannot reach $\Safe$ in at most $k$ steps \-- for much the same reasons {\sc Bounded resilience} and {\sc $k$-resilience} never hold for VAS when $\Safe = \mathop{\downarrow} \Safe$ and $\Bad = \mathop{\uparrow} \Bad$. This lead to a decision procedure
 which lead to the following decidability result:
\begin{theorem}\label{vass down}
{\sc $k$-resilience }  and {\sc Bounded resilience} are decidable for VASS when 
$\Safe = \mathop{\downarrow} \Safe$.
\end{theorem}

\begin{proof}
\iffalse
Let $V$ be a given $d$-VASS and consider a given $k \in \N$.
Consider $\Bad \subseteq Q \times \N^d$ upward-closed and 
$\Safe = S \setminus \Bad$
 downward-closed.
Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition.

@@

From $k_0$ onwards,
$\Bad \setminus (\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))= 
\Bad \setminus (\pred^{*}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$
is stationnary. 
If the set is finite then we stop the procedure once we have checked for every element
whether it is possible to reach $\Safe$ from it.
If it is possible to reach $\Safe$ from every element then {\sc $k_m$-resilience}
hold
for $k_m = \max(k_0, k_{\pi})$ with 
$k_{\pi}$ the maximum of the length of the paths from 
$\Bad \setminus (\pred^{\leq k_0}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))$ to $\Safe$.
If it is infinite then
for every $k \in \N$,
$k$-resilience does not hold, 
and we eventually find a witness that $k$-resilience does not hold.
Indeed, 
if
$\Bad \setminus (\pred^{*}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{*}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$
is
infinite
then
there exists an element of the form $q(\textbf{u} )$
with $\textbf{u} > \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$ 
that is in $\Bad$ but not in any of the $\pred^{\leq k}(\mathop{\uparrow} q_{j_i}(\textbf{0}))$,
and hence from which it is not possible at all to reach
$\Safe$ in $k$ or less steps. 
\fi
\mathieu{Write small text and refer to appendix~\ref{appendix vass}}
\qed
\end{proof}

@@

\section{More details on VASS}\label{appendix vass}

@@

\noindent
\textbf{Theorem~\ref{vass down}.}
{\em {\sc $k$-resilience }  and {\sc Bounded resilience} are decidable for VASS when 
$\Safe = \mathop{\downarrow} \Safe$.
}
